---
title: "PML Course - Assignment Week 4 - Reports"
author: "Gregor Matheis"
date: "March 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

In the context of quantified self movement many devices messure movement constently.
Devices meassured belt, forarm, arm and dumbell movement. The data are used to meassure quality of barbell lifts.

## Preperations

#### Packages

```{r}
library(tidyverse)
library(corrplot)
library(caret)
library(readr)
library(doParallel)
```

#### Data reading

```{r}
pml_training<-read.csv("Data/pml-training.csv",  na.strings = c("NA", "#DIV/0!", ""))
pml_test<-read.csv("Data/pml-testing.csv",  na.strings = c("NA", "#DIV/0!", ""))

```

#### Selecting usable columns

- Excluding date and device informations
- Excluding all variables with missing values

```{r}
pml_training<-pml_training[,!grepl("^X|timestamp|window", names(pml_training))]
pml_test<-pml_test[,!grepl("^X|timestamp|window", names(pml_training))]

classe<-pml_training$classe
pml_training<-pml_training[,sapply(pml_training, is.numeric)]
pml_training$classe<-classe
cols<-apply(pml_training,2, function(x)any(is.na(x)))
cols<-names(cols[!cols])
pml_training<-pml_training[,cols]


problem_id<-pml_test$problem_id
pml_test<-pml_test[,sapply(pml_test, is.numeric)]
pml_test<-pml_test[,cols[1:52]]
pml_test$problem_id<-problem_id
```

#### Subsetting training data into test and train

```{r}
inTrain<-createDataPartition(pml_training$classe, p=.7, list = F)
pml_train<-pml_training[inTrain,]
pml_train_test<-pml_training[-inTrain,]
```

## Data summary

Used variables and their summary
```{r}
data.frame(
"Min" = round(sapply(pml_train[,1:52], min, na.rm=T),2),
"Mean" = round(sapply(pml_train[,1:52], mean, na.rm=T),2),
"SD" = round(sapply(pml_train[,1:52], sd, na.rm=T),2),
"Max" = round(sapply(pml_train[,1:52], max, na.rm=T),2))
```

#### Data correlation

```{r}
correlations<-cor(pml_train[,1:52])
corrplot(correlations, method="circle")
```

## Using random forest

- Within the carret package a random forest can be applied very easiely

- Due to size of dataset i recommend using a parallel version of random forest

#### Setting up model
```{r}
registerDoParallel(cores = 8) #number of cores that is available depents on your laptop
controlRf <- trainControl(method="cv", 5)

model1_rf<-
      train(classe~., data=pml_train,
            method="parRF", trControl=controlRf, 
            importance=TRUE, ntree=100)
```

#### Model sumamry
```{r}
model1_rf
```

#### Model prediction quality
```{r}
model1_rf_pred<-predict(model1_rf, pml_train_test)
confusionMatrix(pml_train_test$classe, model1_rf_pred)
```

#### Interpretation

- The predictors are able to make accurate predictions

- Sensitivity and specificity are very high

- Model is suitable to predict classe variable